---
title: "Lil_Rascals_Lab_11"
author: "Matthew Copeland"
date: "April 3, 2019"
output: html_document
---

```{r}
library(babynames)
library(tidyverse)
library(dbplyr)
library(stringr)
library(forcats)
```


###Matthew's name popularity over time
```{r}
MAT <- babynames %>% filter(name == "Matthew") %>% group_by(year) %>% summarise(n = sum(n))

First3 <- babynames %>% mutate(first3 = substr(name,1,3)) %>% filter(first3 == "Mat")
Sum <- sum(First3$n)
SOY <- First3 %>% group_by(year) %>% summarize(n = sum(n)/Sum)

ggplot()+
  geom_col(data = SOY, aes(x = year, y = n), position = "dodge", fill = "blue")+
  xlab("Year")+
  ylab("Proportion of Names")+
  ggtitle("Proportion of Names That Start With 'Mat'\nThroughout the Years ")

ggplot(data = MAT)+
  geom_col(aes(x = year, y = n), position = "dodge", fill = "red")+
  xlab("Year")+
  ylab("Number of Occurances")+
  ggtitle("The Name 'Matthew' Throughout the Years")
```
Matthew's Findings: Looking at these graphs you can see that the n the first 3 letter "Mat" of my name have been used since the 1800's but looking at the name "Matthew" it seems to start in the 1900's. The trend gradually grows for both graphs as time goes on and then peaks around the late 1900's. The name stayed popular to today and but is on a decline over time.


### James' name popularity over time
```{r}
library(babynames)
library(tidyverse)
library(dbplyr)
library(stringr)
babynames2 <- babynames %>% mutate(first3 = substr(name,1,3)) %>% filter(first3 == "Jam") 
babynames3 <- babynames %>% filter(name == "James") 
Jam <- babynames2 %>% group_by(year) %>% summarise(sum = sum(prop))
James <- babynames3 %>% group_by(year) %>% summarise(sum = sum(prop))
ggplot()+
  geom_col(data = Jam, aes(x = year, y = sum), color = "blue")+
  geom_line(data = James, aes(x = year, y = sum), color = "red")+
  ylab("Proportion of names
       Blue = names that start with Jam
       Red = babies named James")+
  ggtitle("The Name James Throughout the Years")+
  ylab("Year")
```
James Discussion: 
  From the graphs you can see that names starting with Jam increased to its peak from the 20s to about the 50s. It then has quite steadily declined since the 50s until the 20s when it started to plateau. Through the period between 1880 and 2017 the proportion of babies with my name dropped from 0.0502, to 0.0093. However as the name James decreased in popularity, many more similar names were being used

James Summary: 
  I dealt with my individual name and plots of it. I knew that my name was always quite  popular, but I did not know that it was that widespread in the 50s. It is also interesting that although names that start with Jam declined with the name James, not at the same rate, because a lot of similar versions of the name were used.

### Kelsey Individual
```{r}
babynames <- babynames
Kelsey <- babynames %>% mutate(first3 = substr(name,1,3)) %>% filter(first3 == "Kel")
Sum <- sum(Kelsey$n)
yearkel <- Kelsey %>% group_by(year) %>% summarize(n = sum(n)/Sum)
Kelsey <- babynames %>% mutate(first3 = substr(name,1,3)) %>% filter(first3 == "Kel")
babynames1 <- babynames %>% filter(str_detect(name, "Kelsey|Kelsie|Kelsea"))
a1 <- ggplot(data = yearkel)+
  geom_col(aes(x = year, y = n), position = "dodge")+
  geom_vline(xintercept = 1997, color = "red", size = .3)+
  geom_vline(xintercept = 2017, color = "blue", size = .3)+
  xlab("Year")+
  ylab("Proportion of Names")+
  ggtitle("Proportion of Names That Start With 'Kel'\nThroughout the Years ")
a2 <- ggplot(data = babynames1)+
  geom_jitter( mapping = aes(x=year, y = prop, color = sex))+
  geom_smooth( mapping = aes(x=year, y = prop, color = sex))+
    geom_vline(xintercept = 1997, color = "red", size = .3)+
  geom_vline(xintercept = 2017, color = "blue", size = .3)+
  xlab("Year")+
  ylab("Proportion of Names")+
  ggtitle("Proportion of Kelsey Names Throughout the Years ")
plot_grid(a1, a2, ncol=2)

```
Kelsey Discussion: 
  From the graphs you can see that names starting with Kel increased dramatically 
during the 60's. 
In addition, I found that the name Kelsey's popularity increased a lot in the 80's 
then decreased in more recent years. I found that the proportion of Kelsey's in the 
year I was born 1997 was 1.495*10^-2 and in 2017 it was 4.019*10^-3. 
I found that the actual change was a decrease of .010925 and the relative decrease was .73105. I split the gender on the plot in order to deal with the differences in Kel 
names. 


###Ariel names

```{r}
#Different ways to spell ariel
Ariel <- filter(babynames, str_detect(babynames$name, pattern = "(Ariel)|(Arielle)|(Arial)|(Aerial)|(Aurielle)|(Eriel)|(Ariell)|(Erielle)|(Aryel)|(Aeriel)|(Aryelle)|(Arrielle)|(Ariele)|(Aariel)|(Erial)|(Auriel)|(Arieal)|(Aerielle)")) %>% group_by(year) %>% summarise(n = sum(n))
#Names of ariel in three specific years
Ariel %>% filter(year == "1973")
Ariel %>% filter(year == "1988")
Ariel %>% filter(year == "1990")
#probability of someone being named ariel
Ariel2 <- Ariel %>% mutate(prob = (Ariel$n)/133630)
Ariel2 %>% filter(year == "1973")
Ariel2 %>% filter(year == "1988")
Ariel2 %>% filter(year == "1990")
Ariel2 %>% filter(year == "2017")
ggplot(data = Ariel)+
  geom_col(aes(x = year, y = n))+
  ggtitle("Number of Ariel names by Year")+
  xlab("Year")+
  ylab("Number of Occurances")
```

### Rachel Names
```{r}
#Different ways to spell ariel
Rachel <- filter(babynames, str_detect(babynames$name, pattern = "(Rachel)|(Rachael)")) %>% group_by(year) %>% summarise(n = sum(n))
Rachel2 <- filter(babynames, str_detect(babynames$name, pattern = "(Rachel)|(Rachael)")) %>% group_by(year) %>% summarise(prop = sum(prop))
#Names of ariel in three specific years
Rachel %>% filter(year == "1973")
Rachel %>% filter(year == "1988")
Rachel %>% filter(year == "1990")
Rachel %>% filter(year == "2017")
Rachel2 %>% filter(year == "1973")
Rachel2 %>% filter(year == "1988")
Rachel2 %>% filter(year == "1990")
Rachel2 %>% filter(year == "2017")
#graph of ariel names
ggplot(data = Rachel)+
  geom_col(aes(x = year, y = n))+
  ggtitle("Number of Rachel names by Year")+
  xlab("Year")+
  ylab("Number of Occurances")
```
###Ariel and Rachel Summary: 
There were 172 versions of Ariel in 1973, 2167 in 1988, and 6223 in 1990. 
For Rachel there were 8136 in 1973, 19380 in 1988 and 19813 in 1990. 

The proportion of Ariels was .000108 ub 1973, .00112 in 1988, and .00302 in 1990. 
Whereas the proportion of Rachels was .00523 in 1973, .0101 in 1988, .00965 in 1990, and .000953 ub 2017. 
So the proportion of Rachels has always been higher and peaked in 1988ish whereas the proportion of Ariels was always lower but increase steadily over these years. 

The plots show that the name Rachel was unimodal and peaked in 1988ish. The plot of Ariel is bimodal and shows it peaked and the has started gaining popularity again in recent years

###The Little Mermaid Effect
Question: Did The Little Mermaid cause more baby girls to be named Ariel?
```{r}
ggplot(data = Ariel)+
  geom_col(aes(x = year, y = n))+
  geom_vline(xintercept = 1989, color = "red", size = 1)+
  ggtitle("Number of Ariel names by Year")+
  xlab("Year")+
  ylab("Number of Occurances")

ggplot(data = Ariel2)+
  geom_col(aes(x = year, y = prob), position = "dodge")

#graph of female names
FN <- babynames %>% filter(sex == "F") %>% group_by(year) %>% summarise(n = sum(n)) %>% filter(year >= 1988, year <= 1990)
AN <- Ariel %>% filter(year >= 1988, year <= 1990)
FNG <- ggplot(data = FN)+
  geom_smooth(aes(x = year, y = n))
ANG <- ggplot(data = AN)+
  geom_smooth(aes(x = year, y = n))
plot_grid(FNG,ANG, ncol = 2)
```
Findings: As you can see in the graph there is a big spike after 1989 of children being named Ariel as showed by the vertical line. As well the proportion of childeren being named Ariel rises at the 1980 range which would make sense if the number of names goes up then the propbability would go up as well. When looking at the change of all female names and Ariel names they both have the same shape but for all female names the numbers are much higher since its all the names instead of just one



###Team Summary
Kelsey Summary: 
  I dealt with my individual name and plots of it. It was interesting to see how my 
name's popularity peaked in the 80's then has lately become less popular. 
I worked with one data set that combined all the Kelsey's by year then I dealt 
with the Kel only and plotted them both side by side to see the differences and compare. 
I also added the Rachel analysis and stuff for her name.  

Matthew's Summary:
  I worked on my individual name part where I filtered out the names that had the "Mat" and also My full name.
  Once I graphed it I could see the popularity change overtime and made conclusions based on it. I also worked with the Ariel names and the little mermaid where when i graphed you could see the change from when the movie little mermaid came out. 
   I also worked with the Little mermaid part in which i had to manipulate the data to see the change after the movie was released. 